Stack Safety as a Security Property

### 2020-07-10 main.tex
through Fig. 6

@@

      \[\begin{split}
        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
        & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
          \last(\MPT) = \mpostate{\mach_r}{\_}{\_} ~ \land \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
          \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

      Two of the cases of trace confidentiality, \(\mathit{InfCase}\)
      and \(\mathit{FailCase}\), are purely properties of the
      observations produced by the traces they are properties of. If
      the traces are infinite, they should be similar; if one
      fail-stops, it should be a prefix. Only the
      \(\mathit{ReturnCase}\) checks machine states, because in that
      case the traces could continue, but have been cut off by a
      return. The key to making confidentiality lazy is not to perform that
      check but rather to consider the traces after that return,
      checking only whether they are observationally similar.

@@

      {\em Observable trace confidentiality} is the lazy analog to
      trace confidentiality, and reuses the cases in which the traces
      do not return.  For a return predicate \(R\) and a pair of
      traces \(\MPT\) and \(\machT\), we consider the return case, in
      which $\MPT$ has a last element
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and that machine state is
      a return: \(R ~ \mach_r\).
%
      Observable trace confidentiality then requires that $\machT$
      also has a final state \((\mach_r',\obs_r')\) that is a return:
      \(R ~ \mach_r'\).
%
      In addition, just like in the eager case,
      observations of \(\MPT\) are similar to those of \(\machT\).

@@

      \[\begin{split}
        \mathit{ObsReturnCase} ~ R ~
        (\mpostate{\mach_e}{\pol_e}{\obs_e}\MPT) ~
        ((\mach_e',\obs_e')\machT) \ \ \triangleq \ \
        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \Rightarrow\\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mach_r'' \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
        & \pi_o(\MPT') \lesssim \pi_o(\machT') \\
      \end{split}\]

      We combine this case with the other two cases from ordinary trace
      confidentiality:

      \[\begin{split}
        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
        & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
        & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
          \MPT ~ \machT \\
      \end{split}\]

@@

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored
      \[\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r').
      \]
      Addresses 108 and 109
      are not restored; every address that was varied and not changed in one
      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
      from \(\mach_r''\), 108 contains $-4$, and ultimately 38 is printed.
      So observable trace confidentiality does not hold. This example
      illustrates the weakness of observable confidentiality as a testing
      property: it takes until the very end of the program to determine whether
      a tainted value actually influences observable behavior.

>>>>

\definition
\label{def:trace-conf}

@@

      \[\begin{split}
        \propdef{ObsTraceInt} ~ \contour ~ \MPT \ \ \triangleq \ \
        & \head(\MPT) = \mpostate{\mach_e}{\_}{\_} ~ \land ~
          \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
          \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

      Two of the cases of trace confidentiality (\cref{def:trace-conf}), \(\mathit{InfCase}\)
      and \(\mathit{FailCase}\), are defined solely in terms of
      observation traces. If 
      the traces are infinite, they should be similar; if one
      fail-stops, it should be a prefix. Only the
      \(\mathit{ReturnCase}\) checks machine states, because in that
      case the traces could continue, but have been cut off by a
      return. The key to making confidentiality lazy is not to perform that
      check but rather to consider the traces after that return,
      checking only whether they are observationally similar.

@@

      {\em Observable trace confidentiality} is the lazy analog to
      trace confidentiality, and reuses the cases in which the traces
      do not return.  For a return predicate \(R\) and a pair of
      traces \(\MPT\) and \(\machT\), we consider the return case, in
      which $\MPT$ has a last element
      \(\mpostate{\mach_r}{\_}{\_}\) and that machine state is
      a return, i.e. \(R ~ \mach_r\) holds.
%
      Observable trace confidentiality then requires that $\machT$
      also has a final state \((\mach_r',\_)\) that is a return, i.e.
      \(R ~ \mach_r'\) holds.
%
      In addition, just like in the eager case,
      observations of \(\MPT\) are similar to those of \(\machT\).

@@

      \[\begin{split}
        \mathit{ObsReturnCase} ~ R ~
        (\mpostate{\mach_e}{\_}{\_}\MPT) ~
        ((\mach_e',\_)\machT) \ \ \triangleq \ \
        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\_} ~ \land ~
         \last(\machT) = (\mach_r',\_) ~ \land \\
        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') ~ \land \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' ~ \land ~
        \mach_r'' \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT) \eqsim \pi_o(\machT) ~ \land ~
        \pi_o(\MPT') \lesssim \pi_o(\machT') \\
      \end{split}\]

      We combine this case with the other two cases from ordinary trace
      confidentiality:
\[
        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \ \ \triangleq \ \
         \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT ~ \land 
         \mathit{InfCase} ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
          \MPT ~ \machT 
\]

@@

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored
      \[\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r').
      \]
      Addresses 108\apt{107??} and 109\apt{108??}
      are not restored; every address that was varied and not changed in one
      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
      from \(\mach_r''\), 108 contains $-4$, and ultimately 38 is printed.
      So observable trace confidentiality does not hold. This example
      illustrates the weakness of observable confidentiality as a testing
      property: it takes until the very end of the program to determine whether
      a tainted value actually influences observable behavior.
